\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{hyperref}

\title{Peer Evaluation Form for Group Work}
\author{Ian Mooney}
\date{5/6/25}

\begin{document}

\maketitle

\section*{Course: CS3-Intro to Algorithms}
\vspace{0.5cm}

\section*{Project Name : Assignment 4}
\vspace{0.5cm}

\section*{Group Members: Patrick M \& Russell C}
\vspace{1cm}

\section*{Instructions:}

This form is designed to evaluate the contributions of each group member, including yourself, to the project. The evaluations will be used to calculate a multiplier that adjusts each member's final project grade. 

The multiplier will range from \textbf{0.0 (0\% of the grade)} to \textbf{1.0 (full grade)} based on your contributions compared to the group’s highest contributor.

\textbf{Important:} This form must be completed during a group meeting, with all members present. Open discussion and transparency are encouraged to ensure a fair assessment process.

\subsection*{Steps:}

\begin{enumerate}
    \item \textbf{Individual Scoring:} Evaluate yourself and each group member on the following criteria using a scale of 0 to 10 (0 = no contribution, 10 = maximum contribution).
    \item \textbf{Total Score Calculation:} Sum the scores for each criterion to get a total score for each member.
    \item \textbf{Compute the Multiplier:} The multiplier for each member is calculated using the following formula:
    \[
    \text{Multiplier} = \frac{\text{Individual Total Score}}{\text{Highest Total Score in the Group}}
    \]
    This multiplier will be applied to the project grade to determine each member’s final grade.
\end{enumerate}

\section*{Part 1: Self-Evaluation}

\begin{tabular}{|l|c|p{7cm}|}
\hline
\textbf{Criteria} & \textbf{Score (0-10)} & \textbf{Comments (Optional)} \\ \hline
Contribution to Group Discussions & 10 & \\ \hline
Quality of Work & 10 & \\ \hline
Reliability & 10 & \\ \hline
Collaboration \& Support & 10 & \\ \hline
\end{tabular}

\vspace{0.5cm}
\textbf{Total Score (out of 40):} \underline{\hspace{3cm}40}

\section*{Part 2: Peer Evaluation}

\textbf{For each group member, assign a score and provide any comments if necessary.}

\subsection*{Group Member 1: [Russell C.]}

\begin{tabular}{|l|c|p{7cm}|}
\hline
\textbf{Criteria} & \textbf{Score (0-10)} & \textbf{Comments (Optional)} \\ \hline
Contribution to Group Discussions & 10 & \\ \hline
Quality of Work & 10 & \\ \hline
Reliability & 10 & \\ \hline
Collaboration \& Support & 10 & \\ \hline
\end{tabular}

\vspace{0.5cm}
\textbf{Total Score (out of 40):} \underline{\hspace{3cm}40}

\subsection*{Group Member 2: [Patrick M.]}

\begin{tabular}{|l|c|p{7cm}|}
\hline
\textbf{Criteria} & \textbf{Score (0-10)} & \textbf{Comments (Optional)} \\ \hline
Contribution to Group Discussions & 10 & \\ \hline
Quality of Work & 10 & \\ \hline
Reliability & 10 & \\ \hline
Collaboration \& Support & 10 & \\ \hline
\end{tabular}

\vspace{0.5cm}
\textbf{Total Score (out of 40):} \underline{\hspace{3cm}40}

\section*{Final Multiplier Calculation}

After each group member has completed their evaluations, use the following formula to compute the multiplier for each individual:

\[
\text{Multiplier} = \frac{\text{Individual Total Score}}{\text{Highest Total Score in the Group}}
\]

\textbf{Example:} If the highest total score in the group is 38 and your total score is 35, your multiplier would be \(\frac{35}{38} = 0.92\). This means you would receive 92\% of the project grade.

Each group member's final grade for the project will be calculated by multiplying the project grade by their individual multiplier. My multiplier: $40/40 = 100\%$

\section*{Group Discussion Acknowledgment}

By signing below, you confirm that this evaluation was completed collaboratively during a group meeting and that the scores reflect a fair assessment of contributions.

\begin{tabbing}
\hspace{8cm}\=\kill
\underline{\hspace{3.0cm}Ian D. Mooney} \> \textbf{Date:} \underline{\hspace{3cm}5/6/25} \\
\textbf{Signature of Group Member 1} \\
\\
\underline{\hspace{2.5cm}Patrick D. Murphy} \> \textbf{Date:} \underline{\hspace{3cm}5/6/25} \\
\textbf{Signature of Group Member 2} \\
\\
\underline{\hspace{2.5cm}Russell R. Cannon} \> \textbf{Date:} \underline{\hspace{3cm}5/6/25} \\
\textbf{Signature of Group Member 3} \\
\\

\end{tabbing}

\end{document}
